---
title: "Association rule mining"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The .txt file has data of **9,835** rows with each row being a list of items seperated by commas and each row being a new basket.

```{r,echo=FALSE,include=FALSE}
library(ggplot2)
library(tidyverse)
library(arules)
library(arulesViz)

groceries = read.table("groceries.txt",sep="\n",header=FALSE,stringsAsFactors = FALSE)

```

Displaying a sample basket to check the contents and verify if they have been extracted correctly
```{r}
groceries[1,]   
```

Currently our data is just comma separated items with each row of our data being one transaction. 
To use the transaction conversion function we need indivudal baskets as a vector with the items from every transaction.
```{r,include=FALSE}
item_list = list()
```

Splitting items single strings from every basket into individual strings of items and then displaying one row to verify. 
```{r}
item_list = apply(groceries, MARGIN = 1, FUN = function(x){unlist(strsplit(x, ","))})
item_list[1]
```
This is then converted to a vector by **unlisting**.
```{r,echo= FALSE}
item_vector = unlist(item_list)
```
# How many items are present in our dataset?
```{r}
length(unique(item_vector))  
```

There are **169** unique items in our dataset.

```{r,echo=FALSE}
item_count = as.data.frame(table(item_vector))
```

# What is the most popular item?
```{r}
ggplot(item_count %>% 
        arrange(-Freq) %>% 
        head(20), 
        aes(x = reorder(item_vector, -Freq), y = Freq)) + 
  geom_bar(stat = "identity") + 
  coord_flip() 
```

The top 3 most commonly purchased items are **Whole Milk**,**Vegetables** and **Rolls/Buns** in the same order.

# What are the 10 least popular items?
```{r}
item_count %>%
  arrange(Freq) %>%
  head(10)
```

We then convert our items into transactions using the "as" function with "transactions" as the second argument.
```{r,echo=FALSE}
item_list = lapply(item_list, unique)

item_transactions = as(item_list, "transactions")
```

# Applying the Apriori algorithm to identify the most popular associations:
```{r,echo=FALSE,include=FALSE}
groceryrules = apriori(item_transactions,parameter=list(support = .001,confidence = .6))
summary(groceryrules)

inspect(head(groceryrules, n = 10, by ="lift"))
```

The popular associations can be visualised as below:
```{r}
plot(head(groceryrules, n = 10, by ="lift"), method = "graph", main = "Popular association rules")
```

# Conclusions: 
A few of the most common purchases or inferences are:

1. People purchasing **Instant food products** and **soda** are more likely to get **hamburger meat** too.

2. **Cheese, ham, white bread** and **eggs** usually sell together.

3. **Popcorn, soda** are sold popularly with **salty snacks** (possible movie/game night meals).

4. A lot of dairy like **cheese, curd, whipped cream and yogurt** are purchased together.
