---
title: "Portfolio Modelling"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this markdown file we will construct three different portfolios of exchange-traded funds, or ETFs, and use bootstrap resampling to analyze the short-term tail risk of your portfolios. 

```{r, include=FALSE}
library(mosaic)
library(foreach)
library(quantmod)
```

First let us import few stocks. We will choose stocks such that few are volatile and few are stable. Also we chose stocks which have more than 5 years of data.

1. Find stocks to put in your EFT
2. Plot the trends for the past 5 years to see which ones are volatile and which ones are stable
3. Decide the weightes for all 3 portfolios
4. Create a function to calculate the VAR at the end of 4 weeks
Function: 
1. Build a loop to resample the returns day by day and add a rebalancing at the end of the loop
2. Build another loop to perform monte carlo simulations

Stocks were selected from the top 100 most heavily traded exchange products.
```{r echo=TRUE}
mystocks = c("EEM", "QQQ", "SPY", "EFA", "XOP")
getSymbols(mystocks)
```

To adjust for splits and divedends, we use the adjustOLHC package which is part of the quantmod() library.

```{r}
adjusted_stocks = lapply(mystocks, function(i) {
  adjustOHLC(get(i, pos=.GlobalEnv), symbol.name=i, adjust=c("split"), 
             use.Adjusted=FALSE)
})

all_returns = cbind(ClCl(EEM),ClCl(QQQ),ClCl(SPY),ClCl(EFA),ClCl(XOP))
#head(all_returns)
all_returns = as.matrix(na.omit(all_returns))
summary(all_returns)
```

```{r echo=TRUE}
#par(mfrow = c(3, 2))
colours = c('dark red', 'dark cyan','blue','dark green','orange')
for (i in seq_along(mystocks)) {
  df_expr = paste(mystocks[i],"$",mystocks[i],".Close", sep = "")
  data = eval(parse(text = df_expr))
  plot = plot(data, col = colours[i], type = "l", main = paste(mystocks[i],"Closing Prices"))
  print(plot)
}
```


Let us look at the standard deviation to judge the volatility
```{r echo=TRUE}
stdev_stocks = data.frame()
for (i in 1:ncol(all_returns))
{
 s = sd(all_returns[, i], na.rm = T)
 stdev_stocks[1,i] =  s
}
colnames(stdev_stocks) = colnames(all_returns)
stdev_stocks
```

Looks like **EEM** and **XOP** are the most volatile stocks.

Now to build the function to calculate the variance:
```{r echo=TRUE}
days = 20
initial_wealth = 10000
func_wealthtracker = function(weights){
  wealthtracker = matrix(0, nrow = 5000, ncol = days)
  for (i in 1:5000) {
  wealth = initial_wealth
    for (day in 1:20) {
      wealth = initial_wealth
      splits = weights*wealth
      sample.day = resample(all_returns, 1, orig.ids = FALSE)
      return = splits+splits*sample.day
      wealth = sum(return)
      wealthtracker[i,day] <- wealth
    }}
  wealthtracker
}
```

```{r echo=TRUE}
#plot the histogram for the first portfolio
weights1 = c(0.2, 0.2, 0.2, 0.2, 0.2)
data1 = func_wealthtracker(weights1)
ggplot(mapping = aes(data1[,days]-initial_wealth)) +
  geom_histogram(bins = 30, fill = 'black') +
  labs(x = "Profit",
       y = "Frequency",
       title = "VAR - Strategy 1",
       subtitle = "")
quantile(data1[,days] - initial_wealth, 0.05)
```

Thus, from a portfolio where the weights are evenly distributed, we get a VAR of `r quantile(data1[,days] - initial_wealth, 0.05)`. Additionally, we get an average profit of `r round(mean(data1[,days]),0)`.

Now let us see a safer portfolio, where we put  lesser weight on the 2 volatile stocks
```{r echo=TRUE}
weights2 = c(0.05, 0.3, 0.3, 0.3, 0.05)
data2 = func_wealthtracker(weights2)
ggplot(mapping = aes(data2[,days]-initial_wealth)) +
  geom_histogram(bins = 30, fill = 'black') +
  labs(x = "Profit",
       y = "Frequency",
       title = "VAR - Strategy 2",
       subtitle = "")
quantile(data2[,days] - initial_wealth, 0.05)
```

Thus, from a safer portfolio, we get a VAR of **`r quantile(data2[,days] - initial_wealth, 0.05)`**. Additionally, we get an average profit of **`r round(mean(data2[,days]),0)`**.

Next, we will make a more risky portfolio, adding weight on the more volatile stocks
```{r echo=TRUE}
weights3 = c(0.25, 0.2, 0.2, 0.1, 0.25)
data3 = func_wealthtracker(weights2)
ggplot(mapping = aes(data3[,days]-initial_wealth)) +
  geom_histogram(bins = 30, fill = 'black') +
  labs(x = "Profit",
       y = "Frequency",
       title = "VAR - Strategy 3",
       subtitle = "")
quantile(data3[,days] - initial_wealth, 0.05)
```

Thus, from a safer portfolio, we get a VAR of **`r quantile(data3[,days] - initial_wealth, 0.05)`**. Additionally, we get an average profit of **`r round(mean(data3[,days]),0)`**.
